A Horrocks-type theorem for even orthogonal $K_2$

### 2020-04-12 dl-paper.tex
Fix 33-34

\subsection{Proof of Horrocks theorem.} \label{sec:P1glueing}
We retain the notation of the previous section and assume additionally that the map $j_R \colon \St(\Phi, B) \to \St(\Phi, R)$ is injective.
This assumption allows us to consider the group $\St(\Phi, B)$ as a subgroup of $\St(\Phi, R)$.
Being considered in this way, $\St(\Phi, B)$ contains the subgroup $\overline{G}^{\geq 0}_M$ of $\St(\Phi, R)$.
This follows from a consideration of the following commutative diagram, in which the map $t$ is obtained from~\cref{lem:lemma32}:
\[ \xymatrix{ \St(\Phi, A[X], M[X]) \ar[d] \ar[r] & \St(\Phi, R, I) \ar[r]^(.65){t} & \St(\Phi, B) \ar[d]^{j_R} \\
   \St(\Phi, A[X]) \ar[rr]^{j_+} & & \St(\Phi, R). } \]

\begin{lemma}\label{lem:action} For any $[p, h, u]\in \overline{V}_T$ and any $p_1 \in \overline{G}^{\geq 0}_M$ holds $p_1 \cdot [p, h, u] = [p_1p, h, u]$. \end{lemma}
\begin{proof} First of all, notice that the assertion of the lemma can be verified directly in the case $p_1 = z_\alpha(f, a)$ for all $a\in A$, $f\in M[X]$. Using~\cref{lem:Smult} and the fact that $x_\beta(X^2f) \in P_\alpha(0)$ for all $\beta \in \Phi$ it is not hard to show that the assertion also holds for $p_1 = z_\alpha(X^2f, aX^{-1})$.

It suffices to verify the assertion of the lemma for the generators $p_1 = z_\alpha(f, \xi)$ of $\overline{G}_M^{\geq 0}$ for all $f \in M[X]$, $\xi\in A[X]$, $\alpha \in \Phi$. We accomplish this by 
induction on the degree of $\xi$ in $X$. The base of induction is clear.

@@

\begin{comment}
\begin{equation} \label{T_2} t_\alpha(aX^{-1}) (g, h, u) = \left[ x_\alpha(aX^{-1})\cdot g \cdot x_\alpha\left(-\tfrac{aX^{-1}}{1 + am}\right) \{ X, 1+am \} ,\ x_\alpha\left(\tfrac{aX^{-1}}{1 + am}\right) h ,\ u \cdot (1 + am)\right].\end{equation}

\begin{multline*}
 t_\alpha(aX^{-1}) (x_{-\alpha}(mX), 1, 1) = \left[ x_\alpha(aX^{-1})\cdot x_{-\alpha}(mX) \cdot x_\alpha\left(-\tfrac{aX^{-1}}{1 + am}\right) \{ X, 1+am \} ,\ x_\alpha\left(\tfrac{aX^{-1}}{1 + am}\right),\  (1 + am)\right] = \\
 = \left[ x_{-\alpha}\left(\tfrac{mX}{1+am}\right) \cdot \langle a, m\rangle_\alpha \cdot h_\alpha(1+am) ,\ x_\alpha\left(\tfrac{aX^{-1}}{1 + am}\right),\  (1 + am)\right]
\end{multline*}

This implies that 
\begin{multline*} x_{\alpha}\left(\tfrac{-aX^{-1}}{1+am}\right) \cdot h_\alpha(1+am)^{-1} \cdot \langle a, m \rangle^{-1} \cdot x_{-\alpha}\left(\tfrac{-mX}{1+am}\right) x_\alpha(aX^{-1}) \cdot x_{-\alpha}(mX) [1, 1, 1] = [1, 1, 1+am]
\end{multline*}

\begin{equation} \langle a,m \rangle _ \alpha = x_{-\alpha}\left(\tfrac{- m}{1 + am}\right) \cdot x_{\alpha}(a) \cdot x_{-\alpha}(m) \cdot x_{\alpha}\left(\tfrac{- a}{1+am}\right) \cdot h_{\alpha}^{-1}(1 + am). \end{equation} 
\begin{equation} \langle a,m \rangle _ \alpha ^{-1} =  h_{\alpha}(1 + am) \cdot x_{\alpha}\left(\tfrac{a}{1+am}\right) \cdot x_{-\alpha}(-m) \cdot x_{\alpha}(-a) \cdot x_{-\alpha}\left(\tfrac{m}{1 + am}\right). \end{equation} 
\end{comment}

\begin{externaltheorem} \label{thm:P1glueing}
 Let $\Phi$ be a root system of type $\rA_{\geq 4}, \rD_{\geq 5}$ or $\rE_{6,7,8}$.
 Suppose that the map $j_R \colon \St(\Phi, B) \to \St(\Phi, A[X, X\inv])$ is injective.  
 Then the map $j_-$ is injective and the following commutative square is pullback
 \begin{equation} \nonumber
  \xymatrix{ \St(\Phi, A) \ar[r] \ar[d] & \St(\Phi, A[X]) \ar[d]^{j_+} \\ \St(\Phi, A[X\inv]) \ar[r]^{j_-} & \St(\Phi, A[X, X\inv]).}
 \end{equation}
\end{externaltheorem}
\begin{proof} Since $j_- = j_R j_B^-$, where $j_B^-$ is as in~\cref{prop43}, it suffices to show that $j_B^-$ is injective.
 Set $v_0 = [1, 1, 1] \in \overline{V}_T$ and suppose $g \in \Ker(j_B^-)$.
 By~\cref{prop43} holds $v_0 = j_B^-(g) \cdot v_0 = [1, g, 1]$, therefore by the definition of $\overline{V}_T$, $g = i_-(\gamma)$ for some $\gamma \in G_M^0$ such that $j_+i_+(\gamma) = 1$.
 Since $j_+i_+$ is injective, we conclude that $g = 1$.

>>>>

\subsection{Proof of Horrocks theorem.} \label{sec:P1glueing}
We retain the notation of the previous section and assume additionally that the map $j_R \colon \St(\Phi, B) \to \St(\Phi, R)$ is injective.
This assumption allows us to consider the group $\St(\Phi, B)$ as a subgroup of $\St(\Phi, R)$.
Being considered in this way, $\St(\Phi, B)$ contains the subgroup $\overline{G}^{\geq 0}_M$ of $\St(\Phi, R)$.
This follows from a consideration of the following commutative diagram, in which the map $t$ is obtained from~\cref{lem:lemma32}:
\[ \xymatrix{ \St(\Phi, A[X], M[X]) \ar[d] \ar[r] & \St(\Phi, R, I) \ar[r]^{t} & \St(\Phi, B) \ar[d]^{j_R} \\
   \St(\Phi, A[X]) \ar[rr]^{j_+} & & \St(\Phi, R). } \]

\begin{lemma}\label{lem:action} For any $[p, h, u]\in \overline{V}_T$ and any $p_1 \in \overline{G}^{\geq 0}_M$ one has $p_1 \cdot [p, h, u] = [p_1p, h, u]$. \end{lemma}
\begin{proof} 
In view of~\cref{thm:Tits} it suffices to verify the assertion of the lemma for the generators $p_1 = z_\alpha(f, \xi)$ of $\overline{G}_M^{\geq 0}$ for all $f \in M[X]$, $\xi\in A[X]$, $\alpha \in \Phi$. We accomplish this by induction on the degree of $\xi$ in $X$.

First of all, notice that the base case $\xi = a \in A$ can be verified by direct computation involving~\eqref{eq:VVT} and~\eqref{T_leq0}.
Using~\cref{lem:Smult} and the fact that $x_\beta(X^2f) \in P_\alpha(0)$ for all $\beta \in \Phi$ it is not hard to show that the assertion also holds for $p_1 = z_\alpha(X^2f, aX^{-1})$.

@@

\begin{externaltheorem} \label{thm:P1glueing}
 Let $\Phi$ be a root system of type $\rA_{\geq 4}, \rD_{\geq 5}$ or $\rE_{6,7,8}$.
 Suppose that the map $j_R \colon \St(\Phi, B) \to \St(\Phi, A[X, X\inv])$ is injective.  
 Then the map $j_-$ is injective and the following commutative square is pullback
 \begin{equation} \nonumber
  \xymatrix{ \St(\Phi, A) \ar[r] \ar[d] & \St(\Phi, A[X]) \ar[d]^{j_+} \\ \St(\Phi, A[X\inv]) \ar[r]^{j_-} & \St(\Phi, A[X, X\inv]).}
 \end{equation}
\end{externaltheorem}
\begin{proof} Since $j_- = j_R j_B^-$, where $j_B^-$ is as in~\cref{prop43}, it suffices to show that $j_B^-$ is injective.
 Set $v_0 = [1, 1, 1] \in \overline{V}_T$ and suppose $g \in \Ker(j_B^-)$.
 By~\cref{prop43} holds $v_0 = j_B^-(g) \cdot v_0 = [1, g, 1]$, therefore by the definition of $\overline{V}_T$, $g = i_-(\gamma)$ for some $\gamma \in G_M^0$ such that $j_+i_+(\gamma) = 1$.
 Notice that $j_+i_+$ is split by the evaluation homomorphism $ev_{X=a}^*$, $a\neq 0$ and therefore is injective. Thus, we conclude that $g = 1$.
